---
title: "Assignment 5: Text Classification with Naïve Bayes"
author: "Hitesh Bahar"
date: "November 1, 2018"
output: 
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Data Collection from the Guardian API.

```{r}
packages <- c("httr","rjson","jsonlite","dplyr","plyr","stringr","tm","e1071","SnowballC","caret")

lapply(packages, library, character.only = TRUE)
```
* Download Data, Itereatively over six columns in the Guardian API. Append the data into a dataframe **data**
```{r}
api_key<-"d38d159c-3ed5-43b2-9caa-1eed5c665b7e"
url <- "https://content.guardianapis.com/search?"
pageSize<-50
sections<- c("sport", "artanddesign", "business", "culture","technology","world")

data = data.frame()

for (item in sections)
{
  for (i in 1:20)
  {
    req <- fromJSON(paste0(url,("&order-by=newest&show-fields=body&page-size="),pageSize,("&page="),i,("&section="),item,("&api-key="),api_key))
    test <- as.data.frame(req$response$results$fields,flatten = TRUE)
    req$response$results$fields <- NULL
    if("body" %in% colnames(req$response$results) ){
      req$response$results <- rbind(req$response$results,test)
    }else{
      req$response$results <- cbind(req$response$results,test)
    }
    data <- rbind(data,req$response$results)
  }
}
```

* Summarizing the total number of rows collected in each article

```{r}
data %>%
  group_by(data$sectionId) %>%
  dplyr::summarise((count = n()))
```

#2. Data cleaning 
```{r}
# Converting all the column data to lowercase :

columnName <- colnames(data)
for (colN in columnName) {
  data[[colN]] <- tolower(data[[colN]])
}

# To Remove HTML Tags and data cleaning(punctuations,digits) ,


cleanHtmlStringFun <- function(htmlString) {
  return(gsub("<.*?>", "", htmlString))
}

for (i in 1: nrow(data)){
  data$body[i] <- (cleanHtmlStringFun(data$body[i]))
  data$body[i] <- gsub("[[:punct:]]", "", data$body[i] )
  data$body[i] <- gsub(pattern = "\\W", replace = " ",data$body[i])
  data$body[i] <- gsub(pattern = "\\d", replace = " ",data$body[i])
  data$body[i] <- gsub("[[:cntrl:]]", " ", data$body[i] )
  data$body[i] <- str_squish(data$body[i])
  
}

#Printing 3 sample cleaned body data.

print(data[sample(nrow(data), 3), ])
```

#3. Tokenization
```{r}
# Removing Stopwords and single alphabets that are left over 

for (i in 1: nrow(data)){
  data$body[i] <- removeWords(data$body[i],stopwords())
  data$body[i] <- gsub(pattern = "\\b[A-z]\\b{1}", replace = " ", data$body[i])
  data$body[i] <- str_squish(data$body[i])
}

#Extracting Body Column from dataframe Data 

bodyDataCorpus<- VCorpus(VectorSource(data$body))

# Performing Cleaning through Corpus 
bodyDataCorpusClean <- tm_map(bodyDataCorpus,removeNumbers)
bodyDataCorpusClean <- tm_map(bodyDataCorpus,removeWords,stopwords('english'))
bodyDataCorpusClean <- tm_map(bodyDataCorpus,removePunctuation)

# Word Stemming 
bodyDataCorpusClean <- tm_map(bodyDataCorpusClean,stemDocument)

# Creating a DTM  with sparsity oh 97%
bodyDTM <- DocumentTermMatrix(bodyDataCorpusClean)
bodyDTM<- removeSparseTerms(bodyDTM,sparse = 0.97)
bodyDTM<- as.matrix(bodyDTM)
bodyDTM<- as.data.frame(bodyDTM)

print(head(bodyDTM))
```

#4. Classification

```{r}
# Creating Featrure Vector
data$sectionId <- as.factor(data$sectionId)

#Creating Train and Test data  Using 70% Traing and 30% Testing\
bodyDTMTrain <- bodyDTM[1:4200,]
bodyDTMTest<- bodyDTM[4201:6000,]

trainLabel <- data[1:4200,]$sectionId
testLabel <- data[4201:6000,]$sectionId

# Naive Bayes Classifier 
guardian_Classifier <- naiveBayes(bodyDTMTrain,trainLabel)
guardian_Classifier_test <- predict(guardian_Classifier,bodyDTMTest)

#Confusion Matrix 
confusionMatrix(guardian_Classifier_test,testLabel)
```

Here In the above resulst, it is only  predicting from **Class: Technology** and ** Class:World**, As the Y label mostly have this data. Hence we shoould go with random selection of data both on feature vector and the y_label.

```{r}
set.seed(143)
smp_size <- floor(0.75 * nrow(data))
len<- sample(seq_len(nrow(data)), size = smp_size)
bodyDTMTrain <- bodyDTM[len,]
bodyDTMTest<- bodyDTM[-len,]

trainLabel <- data[len,]$sectionId
testLabel <- data[-len,]$sectionId

# Naive Bayes Classifier 
guardian_Classifier <- naiveBayes(bodyDTMTrain,trainLabel)
guardian_Classifier_test <- predict(guardian_Classifier,bodyDTMTest)
```

**Confusion Matrix**
```{r}
#Confusion Matrix 
print(confusionMatrix(guardian_Classifier_test,testLabel))
```

* To improve the performance of  the model, we can do laplace smoothing by *1*. so that the probabilities will atleast not be **0**

```{r}
guardian_Classifier <- naiveBayes(bodyDTMTrain,trainLabel,laplace = 1)
guardian_Classifier_test <- predict(guardian_Classifier,bodyDTMTest)
```

**Confusion Matrix  After Laplace Smoothing**
```{r}
print(confusionMatrix(guardian_Classifier_test,testLabel))
```
